---
title: 'MFA 20222 Pre-programme Assignment'
author: "ELLE LI"
date: "8 September 2021"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    toc: yes
    toc_float: yes
---

```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(gapminder)
library(here)
library(janitor)
```

## Introduction
I am currently a first-year student at London Business School *Masters in Financial Analysis* programme. I am interested in pursuing a Global Markets Summer Analyst Sales and Trading position.\

## Skills
1. I have been **passionate about Financial Markets** since a young age, and I am following market news about stocks, bonds, currencies, derivatives, and commodities daily. I developed huge interests in *New Energy Vehicles* and *Chips* and have done research into these products and sectors during my internships at Genimous and CITIC.
2. My **analytical and quantitative skills** gained from past study and work experience were excellent. During my placement year at UBS, I was responsible for producing monthly risk reports for London branch and EMEA region, providing high level risk profile and analytical commentaries to London branch Chief Risk Officer.
3. Furthermore, with my **cultural background and significant international exposure** both in the UK and China, I believe I would be able to adapt quickly to dynamic work environment and communicate effectively with colleagues and clients. \

## Hobbies and Interests
My life is fast-paced and varied, and I constantly strive to anticipate for dynamic opportunities and aggressively seek for new challenges. Last summer I performed research into what risk the pandemic posed on financial markets, and the paper was published on The Frontiers of Society, Science, and Technology Journal. In addition, I am willing to push beyond what is required. During my placement year at UBS, I have shortened a task from two hours to one second by creating an Excel VBA button and restructured the monthly and quarterly risk reports to better serve managers’ purpose.\

## Link for My LinkedIn Profile
If you are interested for more stories about me, please go to <http://www.linkedin.com/in/elle-li-2000>.

# Task 2: Gapminder dataset

```{r}
glimpse(gapminder)

head(gapminder, 20)

```

I create the `country_data` and `continent_data` with the code below.

```{r}
country_data <- gapminder %>% 
            filter(country == "China")

continent_data <- gapminder %>% 
            filter(continent == "Asia")
```

First, I create a plot of life expectancy over time for *China*, mapping `year` on the x-axis, and `lifeExp` on the y-axis.

```{r, lifeExp_one_country}
plot1<-ggplot(data = country_data, mapping = aes(x = year, y = lifeExp))+
geom_point() +
geom_smooth(se = FALSE)
```

Next I add a title.

```{r, lifeExp_one_country_with_label}
plot1<- plot1 +
labs(title = "Life Expectancy over Time for China",
x = "Year",
y = "Life Expectancy")
```

Secondly, I produce a plot for all countries in *Asia*.

```{r lifeExp_one_continent}
ggplot(continent_data, mapping = aes(x = year, y = lifeExp , colour= country, group = country))+
geom_point() + 
geom_smooth(se = FALSE)
```

Finally, using the original `gapminder` data, I produce a life expectancy over time graph, grouped by continent.

```{r lifeExp_facet_by_continent}
ggplot(data = gapminder , mapping = aes(x = year, y =  lifeExp, colour= country))+
geom_point() + 
geom_smooth(se = FALSE) +
facet_wrap(~continent) +
theme(legend.position="none")
```

Given these trends, it can be seen that life expectancy rose from 1952 to 2007 across all continents, with differences in the patterns.

From the 1960s to 1970s on, the cardiovascular disease mortality declined, leading a decrease in death cases in many developed countries, such as the United States. CVD mortality rate declined in that period mostly since there was a decline in cigarette smoking and tabacco usage. Furthermore, as medical condition advanced, patients having CVD had been well treated and controlled.

Whilst most continents witnessed a steady increase in life expectancy over the 50 years, there were flat or even decreasing lines for most of African countries at around 52 years in 1980s. That was because the HIV or AIDS epidemic quickly became one of the main causes of death in Africa. Thereafter, life expectancy began to rise again at a lower pace, and it was over 64 years in 2020.